---
title: 'Create of SDM'
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---
<<<<<<< HEAD
To create a shiny web page, it is necessary to install the complement
```{r}
install.packages('rsconnect')
```
On secret it is necessary to put the secret key from shiny
```{r}
rsconnect::setAccountInfo(name='je-sh', token='ECB646B152BD92630EE0DDA629AAE6B7', secret='secret')
```
And it upload the page
=======
```{r}
install.packages('rsconnect')
```
Aqui se pone el token de la pagina de shiny
```{r}
rsconnect::setAccountInfo(name='je-sh', token='ECB646B152BD92630EE0DDA629AAE6B7', secret='secret')
```
conecta y sube la pagina a shiny
>>>>>>> a37be8544c5d8e187c16f814cd4612a12e14f31f
```{r}
library(rsconnect)
deployApp()
```
<<<<<<< HEAD
-----------------------------------------------------
=======
>>>>>>> a37be8544c5d8e187c16f814cd4612a12e14f31f
First, we'll download up all the important R packages
```{r}
install.packages('dismo')
install.packages('rgbif')
install.packages('maptools')
install.packages('rJava')

```
Then, we'll load up all the packages
```{r}
library(dismo)
library(rgbif)
library(maptools)
library(rJava)
```
First download on species points contained within mexico. Look up and then insert your invasive species scientific name.
Wesearch for occurrence data on website [GBIF](https://www.gbif.org)
```{r}
<<<<<<< HEAD
your_species_data <- occ_search(scientificName = "Ambystoma mexicanum", 
                          country = 'MX', fields=c('name','country','countryCode','stateProvince','year','decimalLatitude','decimalLongitude'), limit = 20, return ='data') #Limite de informacion
=======
  your_species_data <- occ_search(scientificName = "Ambystoma mexicanum", 
                          country = 'MX', fields=c('name','country','countryCode','stateProvince','year','decimalLatitude','decimalLongitude'), limit = 500, return ='data') #Limite de informacion
>>>>>>> a37be8544c5d8e187c16f814cd4612a12e14f31f
```
Lets make a quick map of the 500 points we downloaded for the species in Mexico.
```{r}
require(leaflet)

xlim <- c(-129,-79) # this tells us to focus on mexico only
ylim <- c(15,35) # this tells us to focus on mexico only

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=your_species_data$data$decimalLongitude, lat=your_species_data$data$decimalLatitude, popup="The birthplace of R")
m
```
### Extracting data

```{r}
path <- file.path(system.file(package="dismo"), 'ex') # This line of code tells R where we stored out climate data.
files <- list.files(path, pattern='grd$', full.names=TRUE )
files # here we see that R found many files related to global climate on the computer.
```


Our next step is to group our climate data.  R calls this "stacking" so now, let's stack up all of our climate data.
```{r}
predictors <- stack(files)
predictors
extent_of_mexico_map <- extent(-129, -79, -15, 35) # Set your extent to the area we're focused on
predictors_cropped_to_mexico <- crop(predictors, extent_of_mexico_map)
```

We can see how the climate data look.  Areas that are colored in green mean those areas have higher temperature or higher precipitation amounts.
```{r}
plot(predictors_cropped_to_mexico)
```
Now, we gather climate information at every blue dot on our map.  This is our way of knowing whether sloths like area with lots of rain, or warm temperatures.  This is the best and most useful part of a species distribution model. We know what conditions species really enjoy.

After taking the next few steps and writing the next lines of code, we have a bunch of data points corresponding to the type of climate that the species likes best, and other data points corresponding to climate the invasive species doesn't like as much.
```{r}
set.seed(0)
group <- kfold(your_species_data$data, 5)
#Creation of data sets to train
pres_train_your_species <- your_species_data$data[group != 1, ]
pres_train_your_species <- as.data.frame(pres_train_your_species[,1:2])

pres_test_your_species <- your_species_data$data[group == 1, ]
pres_test_your_species <- as.data.frame(pres_test_your_species[,1:2])

```
#### What kind of climate does the sloth NOT like?
```{r}
pred_nf <- dropLayer(predictors_cropped_to_mexico, 'biome')
backg <- randomPoints(pred_nf, n=1000, ext=extent_of_mexico_map, extf = 1.25)
```
This bunch of code let us know if the maps we make are accurate.
```{r}
colnames(backg) = c('lon', 'lat')
group <- kfold(backg, 5)
backg_train <- backg[group != 1, ]
backg_test <- backg[group == 1, ]
```
### Maxent
Now, we make our predictive maps using the modeling software called `Maxent`
```{r}
jar <- paste(system.file(package="dismo"), "/java/maxent.jar", sep='')
xm <- maxent(predictors_cropped_to_mexico, pres_train_your_species, factors='biome')
plot(xm)
```
#### Making our maps
There are a few important statistics that tell us if our ecological model is very good.  When we get the maps, the left map is really key. Anywhere we see green, we have a really good chance of seeing sloths.
```{r}
e <- evaluate(pres_test_your_species, backg_test, xm, predictors_cropped_to_mexico)
e
px <- predict(predictors_cropped_to_mexico, xm, ext=extent_of_mexico_map, progress='')
par(mfrow=c(1,2))
plot(px, main='Canis latrans')
require(leaflet)
require(raster)
require(rgdal)
r <- px
pal <- colorNumeric(c("#038247", "#99a10b", "#992806"), values(r), na.color = "transparent")

leaflet() %>% addTiles() %>%
  addRasterImage(r, colors = pal, opacity = 0.6) %>%
  addLegend(pal = pal, values = values(r),
    title = "Surface temp")
```
